---
title: "AE 02: Getting started with RStudio + R Markdown"
subtitle: ""
date: "08.25.2021"
output: 
  html_document: 
    theme: readable
---

## Announcement 

- Professor Tackett will have office hours **Thursday, August 26,9:30a - 11:30a** on Zoom. 
  - Zoom Link: https://duke.zoom.us/j/96752208722
  - Need your Duke NetID & password to join. 
  - Full office hours start August 30.
  
- We'll use the "August 25 Lecture" thread pinned to the top of Ed Discussion: https://edstem.org/us/courses/8027/discussion/

## Questions about the course? 

## Questions from the prepare assignment?

## Tour of R and RStudio 

Let's take a tour of R and RStudio. Go to [vm-manage.oit.duke.edu](https://vm-manage.oit.duke.edu) and open the **STA198-199** Docker containers. 

- Tour of RStudio interface 

- **Console:** Use R as a calculator by typing the following in the console: 

```{r eval = F}
3 * 5 + 10
```

```{r eval = F}
x <- 3
x + x^2
```

```{r eval = F}
x <- 1:10
x*7
```

- Environment
- Loading and viewing a data frame
- Accessing a variable in a data frame
- R functions

## AE-02: Bechdel Test

- Run the following code in the **console** to download the data and Rmd file for today's application exercise. (Note: You will learn how to clone a GitHub repo in tomorrow's lab.) Your RStudio may refresh, and you will see the R Markdown files and data set. 

```{r eval = F}
usethis::use_course("sta199-fa21-003/ae-02")
```

### Understanding the R Project & R Markdown 

- R Project 
- R Markdown file 

### R Markdown vs. Console 

- Run the following in the **console**

```{r eval = F}
x <- 2
x * 3
```

- Then add the following in an R code chunk in your R Markdown file. 

```{r eval = F}
x * 3
```

- What happens? Why the error? 
- Let's fix it! 

### Bechdel test anlaysis 

### GitHub demo 

- Configure git using SSH -- covered in lab tomorrow!
- Clone repo + start new project
- Update document
- `r emo::ji("yarn")` `r emo::ji("white_check_mark")` `r emo::ji("arrow_up")`  knit, commit, push
- Check changes in GitHub repo
- See Commit history


## Submission

**There is nothing to submit for today's AE. Starting Monday, you will submit push your AE to your GitHub repo.**