---
title: "Welcome to MATH 250!"
author: "Dr. Katie Fitzgerald"

output:
  xaringan::moon_reader:
    css: ["xaringan-themer.css", "slides.css"]
    lib_dir: libs
    nature:
      ratio: "16:9"
      highlightLines: true
      highlightStyle: solarized-light
      countIncrementalSlides: false
---

```{r child = "setup.Rmd"}
```

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(rvest)
```

class: center, middle

# Welcome!

---

class: middle, center

# MATH 250: Data Analysis (aka Intro to Data Science)

---

## What is data science?

- Data science is an exciting discipline that allows you to turn raw data into understanding, insight, and knowledge.

- We're going to learn to do this in a `tidy` way -- more on that later!

<!-- - This is a course on introduction to data science, with an emphasis on statistical thinking. -->

---

## Course FAQ

**Q - What data science background does this course assume?**
A - None.

<br>

<!-- **Q - Is this an intro stat course?** -->
<!-- A - While statistics $\neq$ data science, they are very closely related and have tremendous of overlap. Hence, this course is a great way to get started with statistics. However this course is not your typical high school statistics course. -->

<br>

**Q - Will we be doing computing?**
A - Yes. We will use the computing language R.

---

## Course learning objectives

- Learn to explore, visualize, and analyze data in a *reproducible* and *shareable* manner

- Gain experience in data wrangling, exploratory data analysis, predictive modeling, and data visualization

- Work on problems and case studies inspired by and based on *real-world questions* and data

- Learn to effectively communicate results through written assignments and final project presentation

---

## Some of what you will learn

.pull-left[
- Fundamentals of `R`

- Data visualization and wrangling with `ggplot2` and `dplyr` from the `tidyverse`

- Data science workflow and scientific practice

- Web based applications with `RShiny`

- Spatial data visualization

]

.pull-right[
- Data types and functions

- Reproducible reports with `R Markdown`

- Regression and classification

- Statistical inference
]

---

## Examples of data science in practice

- [What redisctricting looks like in every state](https://projects.fivethirtyeight.com/redistricting-2022-maps/)

- [Tracking their life via Fitbit](https://livefreeordichotomize.com/2017/12/27/a-year-as-told-by-fitbit/)

- [US Spread of COVID-19 maps and analytics](https://www.sharedgeo.org/COVID-19/)

---

class: middle, center

# The Course

---

## Course Toolkit

**Course Website:** [apumath250.netlify.app](apumath250.netlify.app)
  - Central hub for the course!
  - Quick tour of the website

**Canvas** [canvas.apu.edu](canvas.apu.edu)
  - Gradebook
  
**Campuswire**: [https://campuswire.com/c/GD02807A2/feed](https://campuswire.com/c/GD02807A2/feed)
  - Class Q&A and discussion forum

**RStudio Cloud:** [https://rstudio.cloud](https://rstudio.cloud)
  - Distribute & work on assignments -- more on this later!

---

## Activities: Prepare, Practice, Perform

- **Prepare**: introduce new content and prepare for lectures
  - Videos & readings before lecture
  - Low-stakes prep quizzes on basic definitions & concepts from videos
  - See *Prepare* column on course schedule 


- **Practice**: practice applying statistical concepts and computing
  - Application Exercises (AE) during Tuesday class, graded for completion

---

## Activities: Prepare, Practice, Perform

- **Perform**: Put together what you've learned to analyze real-world data
  -  Weekly team lab assignments
  -  Periodic individual homework assignments
  -  One take-home exam
  -  Final team project presented during the final exam period.

---

## Statistics Experiences

---

## Where to find help

--

- Attend Office Hours 
  - Tuesdays 10am-12pm
  - Thursdays 1-2:20pm
  - By appointment

--

- Use **Campuswire** for general questions about course content and/or assignments, since other students may benefit from the response. And it's easier to type code and math, compared to email.

--

- Use email for questions regarding personal matters and/or grades.

---

## Course community & Learning environment

---

## Attending class

---

class: middle, center

# Questions? 

**Raise your hand or post on Campuswire**

---

class: middle, center

# Your turn!

---

## Create an RStudio Cloud Account

---

# Let's look at some data!

<!-- - Let's take a look at the [UN Votes](https://apumath250.netlify.app/appex/appex01-unvotes.html) analysis. -->

<!-- - You may discuss with your neighbor. (Don't forget to introduce yourself first!) -->

<!-- - Type your answers under the **August 23 - UN Votes** thread on Ed Discussion -->

<!-- ```{r echo = F} -->
<!-- countdown::countdown(minutes = 5, seconds = 0) -->
<!-- ``` -->

---

## For Thursday





