---
title: "Inference using the Central Limit Theorem"
subtitle: "<br> <br> Introduction to Data Science @ Duke"
author: "[introds.org](https://www.introds.org/)"
output:
  xaringan::moon_reader:
    css: ["xaringan-themer.css", "slides.css"]
    lib_dir: libs
    nature:
      ratio: "16:9"
      highlightLines: true
      highlightStyle: solarized-light
      countIncrementalSlides: false
---

```{r child = "setup.Rmd"}
```

```{r packages, echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(broom)
library(knitr)
library(DT)
library(emo)
library(openintro)
library(infer)
library(patchwork)
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

## The Central Limit Theorem

For a population with a well-defined mean $\mu$ and standard deviation $\sigma$,
these three properties hold for the distribution of sample average $\bar{X}$,
assuming certain conditions hold:

`r emo::ji("white_check_mark")` The distribution of the sample statistic is nearly normal

`r emo::ji("white_check_mark")` The distribution is centered at the (often unknown) population parameter

`r emo::ji("white_check_mark")` The variability of the distribution is inversely proportional to the square 
root of the sample size

---

## Why do we care?

Knowing the distribution of the sample statistic $\bar{X}$ can help us

--

- estimate a population parameter as **point estimate** $\boldsymbol{\pm}$ **margin of error**
  - the .vocab[margin of error] is comprised of a measure of how confident we want to be and how variable the sample statistic is

<br> 
--

- test for a population parameter by evaluating how likely it is to obtain to observed sample statistic when assuming that the null hypothesis is true
  - this probability will depend on how variable the sampling distribution is

---

class: center, middle

## Inference based on the CLT

---

## Inference based on the CLT

If necessary conditions are met, we can also use inference methods based on the CLT. Suppose we know the true population standard deviation, $\sigma$. 

--

Then the CLT tells us that $\bar{X}$ approximately has the distribution $N\left(\mu, \sigma/\sqrt{n}\right)$. 

That is,

$$Z = \frac{\bar{X} - \mu}{\sigma/\sqrt{n}} \sim N(0, 1)$$

---

class: middle

# What if $\sigma$ isn't known?
---

## T distribution

In practice, we never know the true value of $\sigma$, and so we estimate it from our data with $s$.

We can make the following test statistic for testing
a single sample's population mean, which has a .vocab[t-distribution with n-1 degrees of freedom]:

.question[
$$ T = \frac{\bar{X} - \mu}{s/\sqrt{n}} \sim t_{n-1}$$
]

---

## T distribution 


- The t-distribution is also unimodal and symmetric, and is centered at 0

--

- It has thicker tails than the normal distribution
  - This is to make up for additional variability introduced by using $s$ instead of $\sigma$ in calculation of the 
SE

--

- It is defined by the degrees of freedom 
---

## T vs Z distributions

```{r echo=FALSE, out.width = "75%"}
x = seq(-5,5,len=1000)
d = rbind(
  data.frame(dist="t (df=1)",  x=x, d=dt(x,1)),
  data.frame(dist="t (df=3)",  x=x, d=dt(x,3)),
  data.frame(dist="t (df=5)",  x=x, d=dt(x,5)),
  data.frame(dist="t (df=10)", x=x, d=dt(x,10)),
  data.frame(dist="t (df=15)", x=x, d=dt(x,15)),
  data.frame(dist="t (df=30)", x=x, d=dt(x,30)),
  data.frame(dist="Z", x=x, d=dnorm(x))
)

ggplot(d, aes(x=x, y=d, color=dist)) + geom_line()
```

---

## T distribution

.pull-left[
.vocab[Finding probabilities under the t curve:]

```{r}
#P(t < -1.96)
pt(-1.96, df = 9)
```

```{r}
#P(t > -1.96)
pt(-1.96, df = 9, 
   lower.tail = FALSE)
```
]

--

.pull-right[

.vocab[Finding cutoff values under the t curve:]

```{r}
# Find Q1
qt(0.25, df = 9)
```

```{r}
# Q3
qt(0.75, df = 9)
```
]


---

## Resident satisfaction in Durham

`durham_survey` contains resident responses to a survey given by the City of
Durham in 2018. These are a randomly selected, representative sample of
Durham residents.

Questions were rated 1 - 5, with 1 being "highly dissatisfied" and 5 being
"highly satisfied."

--

.question[
Is there evidence that, on average, Durham residents are generally satisfied (score greater than 3) with the quality of the public library system? 
]


---

## Exploratory Data Analysis

.pull-left[
.small[
```{r message=FALSE}
durham <- read_csv("data/durham_survey.csv") %>%
  filter(quality_library != 9)
```
]

.midi[
```{r}
durham %>% 
  summarise(x_bar = mean(quality_library), 
            med = median(quality_library), 
            sd = sd(quality_library), 
            n = n())
```
]
]

.pull-right[

```{r warning = F, echo = F, out.width = "100%"}
durham %>% 
  filter(quality_library!= 9) %>% 
  ggplot(aes(x = quality_library)) + 
    geom_histogram(binwidth = 0.95) +
    labs(x = "Satisfaction with public library system", y = "Count",
         title = "Most residents are generally satisfied <br> with the public library system")
```
]

---

## Hypotheses

.question[
What are the hypotheses for evaluating if Durham residents, on average, are generally satisfied with the public library system?
]

--

$$H_0: \mu = 3$$ 
$$H_a: \mu > 3$$

---

## Conditions

.question[
What conditions must be satisfied to conduct this hypothesis test using methods 
based on the CLT? Are these conditions satisfied?
]

--

**Independence?** 

--

`r emo::ji("white_check_mark")` The residents were randomly selected for the survey, and 521 is less than 10% of the Durham population (~ 270,000).

--

**Sample size / distribution?**

--

`r emo::ji("white_check_mark")` 521 > 30, so the sample is large enough to apply the Central Limit Theorem. 

---

## Calculating the test statistic

Summary statistics from the sample:

```{r fig.height=3, fig.width=5, echo=FALSE}
durham_summary <- durham %>% 
  filter(quality_library != 9) %>%
  summarise(xbar = mean(quality_library), s = sd(quality_library), n = n())
durham_summary
```

--

And the CLT says:

$$\bar{x} \sim N\left(mean = \mu, SE = \frac{\sigma}{\sqrt{n}}\right)$$

--



---

## Calculating the test statistic

.question[
How many standard errors away from the hypothesized population mean is the observed sample 
mean? This is our test statistic. 
]

--

```{r}
(se <- durham_summary$s / sqrt(durham_summary$n)) # SE
```

--

```{r}
(t <- (durham_summary$xbar - 3) / se) # Test statistic
```

---

## Calculating the p-value

.question[
How likely are we to observe a sample mean that is at least as extreme as the 
observed sample mean, if in fact the null hypothesis is true?
]


```{r}
(df <- durham_summary$n - 1) # Degrees of freedom
```

--

```{r}
pt(t, df, lower.tail = FALSE) # P-value, P(T > t |H_0 true)
```

---

## Conclusion

The p-value is very small, so we reject $H_0$.

--

The data provide sufficient evidence at the $\alpha = 0.05$ level that Durham residents, on average, are satisfied with the quality of the public library system $(\mu >3)$

--

.question[
Would you expect a 95% confidence interval to include 3?
]

---

## Confidence interval for a mean

.alert[
**General form of the confidence interval** 

$$point~estimate \pm critical~value \times SE$$
]

--

.alert[
**Confidence interval for the mean** 

$$\bar{x} \pm t^*_{n-1} \times \frac{s}{\sqrt{n}}$$
]

---

## Calculate 95% confidence interval 

.alert[
$$\bar{x} \pm t^*_{n-1} \times \frac{s}{\sqrt{n}}$$
]

--


```{r}
# Critical value 
t_star <- qt(0.975, df)
```

--

```{r}
# Point estimate 
point_est <- durham_summary$xbar
```

--

```{r}
# Confidence interval
CI <- point_est + c(-1,1) * t_star * se
round(CI, 2)
```

---

## Interpret 95% confidence interval 

The 95% confidence interval is `r round(CI[1],2)` to `r round(CI[2],2)`. 

.question[
Interpret this interval in context of the data.
]

--

**We are 95% confident that the true mean rating for Durham residents' satisfaction with the library system is between 3.89 and 4.05.**

---

class: middle

# CLT-based inference using `infer`

---

# CLT-based hypothesis testing in `infer`

$$H_0: \mu = 3 \text{ vs }H_a: \mu > 3$$
--

```{r}
durham %>%
  t_test(response = quality_library, 
         mu = 3, 
         alternative = "greater", 
         conf_int = FALSE)
```

---

## CLT-based confidence intervals in `infer`

Calculate a 95% confidence interval for the mean satisfaction rating.
--

```{r}
durham %>%
  t_test(response = quality_library, 
         alternative = "two-sided",
         conf_int = TRUE, conf_level = 0.95) 
```

---

class: middle

.question[
What is similar, and what is different, between the CLT-based test of means vs. the simulation-based test?
]

